---
title: "CIE chroma (h*) Forest site RF_Regression"
author: "Alex Koiter"
---

## Load libraries
```{r}
suppressPackageStartupMessages({
  library(tidyverse)
  library(randomForest)
  library(terra)
  library(caret)
  library(patchwork)
  library(sf)
})
```

## Load data
```{r}
 attribute <- c("plan_curvature", "profile_curvature", "saga_wetness_index", "catchment_area", "relative_slope_position", "channel_network_distance", "elevation")

data <- read_csv(here::here("./notebooks/forest_terrain_data.csv"), show_col_types = FALSE) %>%
  select("x", "y", "h_c", any_of(attribute))
```

## Map soil property
```{r}

temp_rast <- rast(data)
crs(x = temp_rast, warn=FALSE) <- "epsg:26914"

coords <- read_csv(here::here("./notebooks/coords.csv"), show_col_types = FALSE) %>% 
  st_as_sf(coords = c("long", "lat"),  crs = 4326) %>%
  st_transform(crs = 26914)

p1<- ggplot() +
  tidyterra::geom_spatraster(data = temp_rast, aes(fill = h_c)) +
  scale_fill_viridis_c(name = "h*", breaks = seq(4, 9, 1)) +
  geom_sf(data = filter(coords, site == "Forest")) +
  theme_bw(base_size = 12) +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "bottom") +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +
  ggspatial::annotation_scale(location = "bl") +
  ggspatial::annotation_north_arrow(location = "br")

p2 <- ggplot() +
  tidyterra::geom_spatraster(data = temp_rast, aes(fill = elevation)) +
  scale_fill_viridis_c(name = "Elevation (m)", option = "inferno") +
  geom_sf(data = filter(coords, site == "Forest")) +
  theme_bw(base_size = 12) +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "bottom") +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +
  ggspatial::annotation_scale(location = "bl") +
  ggspatial::annotation_north_arrow(location = "br")
```

```{r}
#| fig-width: 12
#| fig-asp: 0.5
p1+p2  
```

## Create training, validation and testing datasets

60 % of the data in the training data set 20 % of the data in the validation data set 20 % of the data in the testing data set

```{r}
set.seed(123) # makes it reproducible

temp2 <- data %>%
  mutate(dataset = sample(c("train", "validation", "test"), size = nrow(.), replace = TRUE, prob = c(0.6, 0.2, 0.2)))

train <- temp2 %>% 
  filter(dataset == "train" | dataset == "validation") 

validation <- temp2 %>% 
  filter(dataset == "validation") %>%
  select(-dataset)

test <- temp2 %>% 
  filter(dataset == "test") %>%
  select(-dataset)
```

## Feature selection
vifstep() calculates VIF for all variables, excludes the one with the highest VIF (if it is greater than the threshold), repeat the procedure until no variables with a VIF greater than th remains.
```{r}
features <- usdm::vifstep(rast(select(filter(train, dataset == "train"), -dataset, -h_c)), th = 8)
features
```

### Remove correlated features from the data sets
```{r}
train <- temp2 %>% 
  filter(dataset == "train" | dataset == "validation") %>%
  usdm::exclude(features) %>%
  select(-x, -y)

validation <- temp2 %>% 
  filter(dataset == "validation") %>%
  select(-dataset) %>%
  usdm::exclude(features) %>%
  select(-x, -y)

test <- temp2 %>% 
  filter(dataset == "test") %>%
  select(-dataset) %>%
  usdm::exclude(features) %>%
  select(-x, -y)
```

## Tune the training RF model using the validation dataset

Instructions (https://stackoverflow.com/questions/18155482/how-to-specify-a-validation-holdout-set-to-caret) 
This uses the caret package and I included the validation set inside my training set and just define the resampling measures to only use the validation data. This step is to optimize the mtry parameter

```{r}
tc <- trainControl(method = "cv", number = 1, index = list(Fold1 = which(train$dataset == "train")), savePredictions = T)

set.seed(456)
validate.rf <- train(h_c ~ ., data = select(train, -dataset), method = "rf", trControl = tc)
plot(validate.rf)

my_mtry <- validate.rf$finalModel$mtry
my_mtry
```

## Validation back test

Uses the validation dataset as the test
```{r}
set.seed(789)

rf.fit <- randomForest(h_c ~ ., data = select(filter(train, dataset == "train"), -dataset), 
                       ntree = 500, keep.forest = TRUE, importance = TRUE, mtry = my_mtry,
                       ytest = select(filter(train, dataset == "validation"), -dataset)$h_c,
                       xtest = dplyr::select(select(filter(train, dataset == "validation"), -dataset), -h_c))
rf.fit

```

```{r}
oob_val <- sqrt(rf.fit$mse)
test_val <- sqrt(rf.fit$test$mse)

val_plot <- tibble(`Out of Bag Error` = oob_val,
                   Validation = test_val,
                   ntrees = 1:rf.fit$ntree) %>%
  pivot_longer(cols = -ntrees, names_to = "Metric", values_to = "RMSE" )

ggplot(data = val_plot, aes(ntrees, RMSE, color = Metric)) +
  geom_line() +
  theme_bw() +
  xlab("Number of trees")
```

## Testing the RF model

Uses the RF model to predict the test dataset. We compare predicted against actual
```{r}
prediction <- predict(rf.fit, newdata = test)
test_plot <- data.frame(pred = prediction, obs = test$h_c)

r_sq <- summary(lm(pred~obs, data = test_plot))
r_sq

ggplot(data = test_plot, aes(y = pred, x = obs)) +
  geom_point() +
  ggpmisc::stat_poly_line() +
  ggpmisc::stat_poly_eq() +
  theme_bw() +
  geom_abline(slope = 1, intercept = 0) +
  coord_fixed(ratio = 1)
```

## Importance
```{r}
ImpData <- as.data.frame(importance(rf.fit)) %>%
  mutate(Var.Names = row.names(.)) %>%
 `row.names<-`(as.character(1:nrow(importance(rf.fit))))
```
### Top based on IncNodePurity
```{r}
slice_max(ImpData,order_by = IncNodePurity, n = 10)
```

```{r}
p1 <- ggplot(slice_max(ImpData,order_by = IncNodePurity, n = 10), aes(x = fct_reorder(Var.Names, IncNodePurity), y = IncNodePurity)) +
  geom_segment(aes(x = fct_reorder(Var.Names, IncNodePurity), xend = fct_reorder(Var.Names, IncNodePurity), y = 0, yend = IncNodePurity), color="skyblue") +
  geom_point(aes(size = `%IncMSE`), color = "blue", alpha = 0.6) +
  theme_bw() +
  coord_flip() +
  labs(x = "Terrain Attribute") +
  theme(legend.position="bottom")
```

### Top 10 based on %IncMSE
```{r}
slice_max(ImpData,order_by = `%IncMSE`, n = 10)
```

```{r}
p2<- ggplot(slice_max(ImpData,order_by = `%IncMSE`, n = 10), aes(x = fct_reorder(Var.Names, `%IncMSE`), y = `%IncMSE`)) +
  geom_segment(aes(x = fct_reorder(Var.Names, `%IncMSE`), xend = fct_reorder(Var.Names, `%IncMSE`), y = 0, yend = `%IncMSE`), color="skyblue") +
  geom_point(aes(size = IncNodePurity), color = "blue", alpha = 0.6) +
  theme_bw() +
  coord_flip() +
  labs(x = "Terrain Attribute") +
  theme(legend.position="bottom")
```

### Combined

```{r}
p1+p2
```

### Write data

```{r}

importance_data <- ImpData %>%
  mutate(MSE_rank = rank(-`%IncMSE`)) %>%
  mutate(Purity_rank = rank(-IncNodePurity)) %>%
  mutate(site = "Forest",
        property = "h_col")

if(file.exists(here::here("./notebooks/importance_data.csv"))) {
  importance_data_final <- read_csv(here::here("./notebooks/importance_data.csv")) |>
    rows_upsert(importance_data, by = c("MSE_rank", "site", "property")) 
} else importance_data_final <- importance_data

write_csv(importance_data_final, here::here("./notebooks/importance_data.csv"))


model_performance <- tibble(MSE = rf.fit$mse[length(rf.fit$mse)],
       Var_exp = rf.fit$rsq[length(rf.fit$rsq)],
       MSE_test = rf.fit$test$mse[length(rf.fit$test$mse)],
       Var_exp_test = rf.fit$test$rsq[length(rf.fit$test$rsq)],
       R2 = r_sq$r.squared, 
       site = "Forest",
       property = "h_col")

if(file.exists(here::here("./notebooks/model_performance_data.csv"))) {
  model_performance_final <- read_csv(here::here("./notebooks/model_performance_data.csv")) |>
    rows_upsert(model_performance, by = c("site", "property"))  
} else model_performance_final <- model_performance


write_csv(model_performance_final, here::here("./notebooks/model_performance_data.csv"))

```
